---
title: "Madagascar"
subtitle: "Early Warning"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    self_contained: true
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r input, include=FALSE}
# ============================================================
# User inputs
# ============================================================
knitr::opts_chunk$set(
  error = TRUE
)

# Adjust these two values for each country/report
country_name <- "Madagascar"
config_file  <- "MDG_configuration.xlsx"
module <- "ews"   # ews

# Main project path
path <- file.path("..", country_name)

```

```{=html}
<!-- This section imports custom fonts (Inter and Poppins) from Google Fonts
     and loads loads an external CSS stylesheet from the CEA GitHub repository.
     These styles help ensure that the document follows a clean and cohesive
     visual design. CSS style is in CEA github-->
     
     
<link rel="stylesheet" href="https://fonts.googleapis.com/css?family=Inter">
<link rel="stylesheet" href="https://fonts.googleapis.com/css?family=Poppins">
<link rel="stylesheet" href="https://cea-dataportal.github.io/trust-index/asset/style.css">
```

```{r setup_data, include=FALSE}
# ============================================================
# Setup
# ============================================================

# repo <- "https://raw.githubusercontent.com/CEA-dataportal/trust-index/main" #DO NOT CHANGE
repo <- "/Users/vincentturmine/Documents/GitHub/trust-index"

# Module selection

module <- tolower(module)
supported_modules <- c(
  "inst",
  "ews"
)

if (!module %in% supported_modules) {
  stop("Unsupported module: ",module,". Expected: ",paste(supported_modules, collapse = ", "))
}


# Module configuration

source(paste0(repo,"/R/modules/palette_",module,".R")) # Style and colours
source(paste0(repo,"/R/modules/config_",module,".R")) # Module parameters and prefixes


# Core configuration

source(paste0(repo, "/R/base/setup.R")) 
source(paste0(repo, "/R/base/read_config.R"), local = knitr::knit_global())  # from Configuration file to R

cti_config <- read_cti_config(config_file = config_file, country_name = country_name) 
list2env(cti_config, envir = knitr::knit_global())

source(paste0(repo, "/R/base/load_data.R")) # Load data
translation <- "EN"   # ews
source(paste0(repo, "/R/base/translation_v2.R")) # Load CTI translation


```

```{r setup_analysis, include=FALSE}

# Analysis and Score calculation

source(paste0(repo, "/R/base/analysis.R"))

```

::: input-required
*Text to be modified here. / Texte à modifier ici. / Texto a modificar
aquí.*
:::

*Overall perceptions of the Burundian Red Cross are highly positive,
with consistently strong ratings across both competencies and values.
Capability and respectfulness are the highest-rated dimensions, while
effectiveness, engagement, and transparency score comparatively lower
but remain positive. Trust is broadly consistent across provinces and
demographic groups, with only modest variation. In contrast, respondents
with previous engagement with the Burundian Red Cross including
volunteers, beneficiaries, donors, and those who have interacted with
its services or activities consistently report higher levels of trust
than those without prior engagement.*

**`r tr("report.insights")`**

::: input-required
*Text to be modified here. / Texte à modifier ici. / Texto a modificar
aquí.*
:::

- High overall trust: Respondents report consistently positive
  perceptions of the Burundian Red Cross across all competency and value
  dimensions.

- Limited demographic differences: Trust levels remain broadly similar
  across provinces, age groups, gender, and education, indicating that
  positive perceptions are widely shared.

- Engagement strengthens trust: Previous interaction with the Burundian
  Red Cross is associated with higher ratings across both competencies
  and values, with the largest differences observed between respondents
  who have and have not engaged with the organization.

# `r tr("report.sampling")` {.tabset .tabset-fade .tabset-pills}

The survey was conducted across four regions of Madagascar:
Analanjirofo, Atsimo-Andrefana, Menabe and Vatovavy Fitovinany. A
stratified sampling approach was used to achieve coverage across the
study regions and key population groups.

- **Geographic coverage:** The sample closely matched the population
  distribution in Atsimo-Andrefana (33.9% of respondents, compared with
  35.0% of the population). Analanjirofo (32.5% vs. 23.8%) and Menabe
  (17.0% vs. 14.1%) were overrepresented, while Vatovavy Fitovinany was
  underrepresented (16.6% vs. 27.1%).

- **Age and sex:** The overall sex distribution was broadly aligned with
  population estimates (52.7% women and 47.3% men in the sample).
  However, the sample was older than the population: respondents aged
  18–29 accounted for 21.9% of the sample, compared with 42.0% of the
  population. All age groups aged 30 and above were overrepresented,
  particularly those aged 60+ (14.7% vs. 8.7%).

- **Education, employment and disability:** Comparable population data
  were not available for these characteristics, so representativeness
  could not be assessed. Within the sample, 41.3% of respondents
  reported primary education, 31.5% secondary education, 25.1% no formal
  education and 2.2% university education. Regarding employment, 32.6%
  reported informal or irregular work, 26.9% were unemployed and 21.7%
  had regular or part-time employment. In addition, 17.4% of respondents
  reported having a disability.

```{r setup_charts, results='asis'}

source(paste0(repo, "/R/base/charts_sampling.R"))
source(paste0(repo, "/R/base/charts_drivers.R"))
source(paste0(repo, "/R/base/charts_score.R"))
source(paste0(repo, "/R/base/charts_advanced.R"))
source(paste0(repo, "/R/base/tables.R"))


```

```{r geographic_header, results='asis', echo=FALSE}
if (isTRUE(check_adm)) {
  cat("## ", tr("report.geographic_distribution"), "\n\n")
}
```

```{r geographical distribution, echo=FALSE,fig.width=12, fig.height=6, eval=check_adm}

grid.arrange(geo_plot_survey, geo_plot_population, nrow = 1)

```

```{r maps_header, results='asis', echo=FALSE}
if (isTRUE(check_adm)) {
  cat("## ", tr("report.coverage"), "\n\n")
}
```

```{r maps, echo=FALSE, fig.width=8, warning = FALSE, results='asis', eval=check_adm}

stopifnot(exists("map_region"))
stopifnot(inherits(map_region, "tmap"))

map_widget <- tmap::tmap_leaflet(map_region)

htmltools::browsable(
  htmltools::tagList(map_widget)
)

```

## `r tr("report.gender")`

```{r gender_charts, echo=FALSE,fig.width=12, fig.height=7}

pyramid_plot

```

`r if(check_strata1) paste0("## ", tr_variable(strata1_tab))`

```{r strata1, echo=FALSE,fig.width=12, fig.height=5}

## Display a grid with 2 charts
if(check_strata1) {

             if (has_pop_strata1) {
               gridExtra::grid.arrange(
                strata1_plot_survey,
                strata1_plot_population,
                nrow = 1)
             } else {
               gridExtra::grid.arrange(
                strata1_plot_survey,
                placeholder_plot1,
                nrow = 1)
               }
}

```

`r if(check_strata2) paste0("## ", tr_variable(strata2_tab))`

```{r strata2, echo=FALSE,fig.width=12, fig.height=5}

## Display a grid with 2 charts
if(check_strata2) {

             if (has_pop_strata2) {
               gridExtra::grid.arrange(
                strata2_plot_survey,
                strata2_plot_population,
                nrow = 1)
             } else {
               gridExtra::grid.arrange(
                strata2_plot_survey,
                placeholder_plot2,
                nrow = 1)
               }
}


```

`r if(check_strata3) paste0("## ", tr_variable(strata3_tab))`

```{r strata3, echo=FALSE,fig.width=12, fig.height=5}

## Display a grid with 2 charts
if(check_strata3) {

             if (has_pop_strata3) {
               gridExtra::grid.arrange(
                strata3_plot_survey,
                strata3_plot_population,
                nrow = 1)
             } else {
               gridExtra::grid.arrange(
                strata3_plot_survey,
                placeholder_plot3,
                nrow = 1)
               }
}


```

## `r tr("report.limitations")`

::: input-required
*Add a text describing the study limitations. / Ajouter un texte
décrivant les limites de l’étude. / Añadir un texto que describa las
limitaciones del estudio.*
:::

# `r tr("report.results")`

```{r results_intro, results='asis'}

include_text("responses_intro")

```

## `r tr("report.drivers")` {.tabset .tabset-fade .tabset-pills}

`r if(module == "inst") paste0("### ", tr("score.competencies"))`

```{r drivers_competencies, echo=FALSE, fig.width=12, fig.height=10, warning=FALSE, message=FALSE, eval=module == "inst"}

 if (module == "inst") {
  drivers_comp_plot
}
```

`r if(module == "inst") paste0("### ", tr("score.values"))`

```{r drivers_values, echo=FALSE, fig.width=12, fig.height=10, warning=FALSE, message=FALSE, eval=module == "inst"}

if (module == "inst") {
  drivers_val_plot
}


```

`r if(module == "ews" && check_pillar1) paste0("### ", tr("score.pillar1.label"))`

```{r drivers_ews_pillar1, echo=FALSE, fig.width=12, fig.height=10, warning = FALSE,message = FALSE, eval=check_pillar1}

if (module == "ews") {
  drivers_disaster_plot
}
  

```

`r if(module == "ews" && check_pillar2)  paste0("### ", tr("score.pillar2.label"))`

```{r drivers_ews_pillar2, echo=FALSE, fig.width=12, fig.height=10, warning = FALSE,message = FALSE, eval=check_pillar2}

if (module == "ews") {
  drivers_detection_plot
}

```

`r if(module == "ews" && check_pillar3)  paste0("### ", tr("score.pillar3.label"))`

```{r drivers_ews_pillar3, echo=FALSE, fig.width=12, fig.height=10, warning = FALSE,message = FALSE, eval=check_pillar3}

if (module == "ews") {
  drivers_dissemination_plot
}
  

```

`r if(module == "ews" && check_pillar4)  paste0("### ", tr("score.pillar4.label"))`

```{r drivers_ews_pillar4, echo=FALSE, fig.width=12, fig.height=10, warning = FALSE,message = FALSE, eval=check_pillar4}

if (module == "ews") {
  drivers_response_plot
}

  

```

`r if(check_extra) {"### Additional Questions {.tabset .tabset-fade .tabset-pills}"}`

```{r extra_charts, echo=FALSE, results='asis', fig.width=12, fig.height=8}

if (isTRUE(check_extra)) {
  
  for (i in seq_along(extra_plots)) {
    
    cat("#### ", names(extra_plots)[i], "\n\n", sep = "")
    
    print(extra_plots[[i]])
    
    cat("\n\n")
  }
}
```

## `r tr("report.contextual_questions")` {.tabset}

```{r experience_intro, results='asis'}

include_text("experience_intro")

```

`r if (check_experience) paste0("### ", tr("report.exp_questions")) else ""`

```{r experience_question, fig.width=12, fig.height=height_exp_plot, echo=FALSE, warning = FALSE, message = FALSE}

if ("experience" %in% question_code$category) {
  experience_plot
}
```

`r if (check_behaviours) paste0("### ", tr("report.behaviours_questions")) else ""`

```{r behaviours_questions, fig.width=12, fig.height=height_behaviours_plot, echo=FALSE,warning = FALSE,message = FALSE}

if ("behaviours" %in% question_code$category) {
behaviours_plot
    }
```

`r if (check_impact) paste0("### ", tr("report.impact_questions")) else ""`

```{r impact_questions, fig.width=12, fig.height=height_impact_plot, echo=FALSE,warning = FALSE,message = FALSE}

if ("impact" %in% question_code$category) {
impact_plot
    }
```

`r if (check_intention) paste0("### ", tr("report.intention_questions")) else ""`

```{r intention_questions, echo=FALSE, fig.width=12, fig.height=height_intention_plot ,message=FALSE, warning=FALSE}

if("intention" %in% question_code$category) {
intention_plot
    }
```

`r if (check_channel) paste0("### ", tr("report.channel_questions")) else ""`

```{r channel_questions, echo=FALSE, fig.width=12, fig.height=7 ,message=FALSE, warning=FALSE}
if(check_channel) {
plots_by_channel
  }
```

`r if (check_knowledge) paste0("### ", tr("report.knowledge_questions")) else ""`

```{r knowledge_questions, echo=FALSE, fig.width=12, fig.height=7 ,message=FALSE, warning=FALSE}

if(check_knowledge) {
plots_knowledge
  }

```

`r if (check_risk || check_ews) paste0("## ", tr("report.ews")," {.tabset}") else ""`

`r if (check_risk) paste0("### ", tr("report.risk_questions")) else ""`

```{r risk_perception, echo=FALSE, fig.width=12, fig.height=height_risk_plot, warning=FALSE, message=FALSE, eval=check_risk}

if (check_risk){
  risk_grid
}

  

```

`r if (check_ews) paste0("### ", tr("report.ews_questions")) else ""`

```{r ews_questions, echo=FALSE, fig.width=12, fig.height=7 ,message=FALSE, warning=FALSE}

if(check_ews) {
ews_plot
  }

```

# `r tr("report.score")`

```{r  score_intro, echo=FALSE, results='asis'}

include_text("score_intro")

```

## `r tr("report.score_overall")` {.tabset .tabset-fade .tabset-pills}

::: input-required
*Text to be modified here. / Texte à modifier ici. / Texto a modificar
aquí.*
:::

The Burundian Red Cross receives consistently high ratings across all
competency and value dimensions. Capability and respectfulness are the
strongest-performing areas, while engagement and transparency present
the greatest opportunities for improvement.

`r if(module == "ews") paste0("### ", tr("score.bypillar"))`

```{r overall_scores, results='asis', fig.width=12, fig.height=7}

if (module == "inst") {

  gridExtra::grid.arrange(
    comp_plot,
    values_plot,
    ncol = 2
  )

} else if (module == "ews") {

  pillar_plots <- list()

  if (isTRUE(check_pillar1) && exists("pillar1_plot") && !is.null(pillar1_plot)) {
    pillar_plots <- c(pillar_plots, list(pillar1_plot))
  }

  if (isTRUE(check_pillar2) && exists("pillar2_plot") && !is.null(pillar2_plot)) {
    pillar_plots <- c(pillar_plots, list(pillar2_plot))
  }

  if (isTRUE(check_pillar3) && exists("pillar3_plot") && !is.null(pillar3_plot)) {
    pillar_plots <- c(pillar_plots, list(pillar3_plot))
  }

  if (isTRUE(check_pillar4) && exists("pillar4_plot") && !is.null(pillar4_plot)) {
    pillar_plots <- c(pillar_plots, list(pillar4_plot))
  }

  if (length(pillar_plots) > 0) {
    gridExtra::grid.arrange(
      grobs = pillar_plots,
      ncol = 2
    )
  }
}


```

`r if(module == "ews") paste0("### ", tr("score.combined"))`

```{r combined_pillar_score, echo=FALSE, fig.width=12, fig.height=7, warning=FALSE, message=FALSE}

if (module == "ews") {
  combined_pillar_plot
}
```

## `r tr("report.score_factors")`{.tabset .tabset-fade .tabset-pills}

::: input-required
*Text to be modified here. / Texte à modifier ici. / Texto a modificar
aquí.*
:::

Perceptions of the Burundian Red Cross are consistently positive across
demographic, geographic, and engagement groups. Competency and value
scores are generally high, with only modest variation between subgroups,
indicating a strong and broadly shared level of trust across the
surveyed population.

*Distribution of mean scores for values and competencies per demographic
questions*

```{r factor_scores, results='asis', fig.width=12, fig.height=height_strip_plot}

if (module == "inst") {
  grid.arrange(strip, co, va,ncol = 3,widths = c(1, 4, 4)
  )
}

```

`r if(module == "ews" && check_pillar12) paste0("### ", tr("score.pillar12"))`

```{r factor_scores_ews12, class.source="next", echo=FALSE,fig.width=15, fig.height=19,warning = FALSE,message = FALSE,results = "hide", eval=check_pillar12}

if (module == "ews"){
plots <- list(strip)

widths <- c(1)

if (isTRUE(check_pillar1) && !is.null(p1_factor_plot)) {
  plots <- c(plots, list(p1_factor_plot))
  widths <- c(widths, 4)
}

if (isTRUE(check_pillar2) && !is.null(p2_factor_plot)) {
  plots <- c(plots, list(p2_factor_plot))
  widths <- c(widths, 4)
}

gridExtra::grid.arrange(
  grobs = plots,
  ncol = length(plots),
  widths = widths
)

}
```

`r if(module == "ews" && check_pillar34) paste0("### ", tr("score.pillar34"))`

```{r factor_scores_ews34, class.source="next", echo=FALSE,fig.width=15, fig.height=19,warning = FALSE,message = FALSE,results = "hide", eval=check_pillar34}

if (module == "ews") {
plots <- list(strip)

widths <- c(1)

if (isTRUE(check_pillar3) && !is.null(p3_factor_plot)) {
  plots <- c(plots, list(p3_factor_plot))
  widths <- c(widths, 4)
}

if (isTRUE(check_pillar4) && !is.null(p4_factor_plot)) {
  plots <- c(plots, list(p4_factor_plot))
  widths <- c(widths, 4)
}

gridExtra::grid.arrange(
  grobs = plots,
  ncol = length(plots),
  widths = widths
)
}
```

## `r tr("report.score_profile")`

::: input-required
*Text to be modified here. / Texte à modifier ici. / Texto a modificar
aquí.*
:::

Competency and value perceptions remain consistently high across
engagement groups. Respondents who have volunteered with or previously
benefited from the Burundian Red Cross tend to rate the organization
slightly more positively than those without prior engagement, although
trust remains strong across all groups.

```{r profile_scores, results='asis', fig.width=12, fig.height=8 }

if (module == "inst") {
  grid.arrange(c_plot, v_plot, ncol = 2)
}



```

`r if(module == "ews" && check_pillar12) paste0("### ", tr("score.pillar12"))`

```{r sscoring_profile1, echo=FALSE,fig.width=12, fig.height=8,warning = FALSE,message = FALSE,results = "hide",eval=check_pillar12}


if (module == "ews") {
plots <- list()

if (isTRUE(check_pillar1) && !is.null(p1_profile_plot)) {
  plots <- c(plots, list(p1_profile_plot))
}

if (isTRUE(check_pillar2) && !is.null(p2_profile_plot)) {
  plots <- c(plots, list(p2_profile_plot))
}

if (length(plots) > 0) {
  gridExtra::grid.arrange(
    grobs = plots,
    ncol = min(2, length(plots))
  )
}
}
```

`r if(module == "ews" && check_pillar34) paste0("### ", tr("score.pillar34"))`

```{r scoring_profile2, echo=FALSE,fig.width=12, fig.height=8,warning = FALSE,message = FALSE,results = "hide",eval=check_pillar34}

if (module == "ews") {
plots <- list()

if (isTRUE(check_pillar3) && !is.null(p3_profile_plot)) {
  plots <- c(plots, list(p3_profile_plot))
}

if (isTRUE(check_pillar4) && !is.null(p4_profile_plot)) {
  plots <- c(plots, list(p4_profile_plot))
}

if (length(plots) > 0) {
  gridExtra::grid.arrange(
    grobs = plots,
    ncol = min(2, length(plots))
  )
}
}

```

# `r tr("report.methods_metrics")` {.tabset .tabset-fade .tabset-pills}

## `r tr("report.methods")`

```{r methodology, echo=FALSE, results='asis'}

include_text("methods")

```

## `r tr("report.metrics")`

```{r tables, results='asis'}
# ============================================================
# Tables
# ============================================================

cat("### ", tr("table.gender"), "\n\n")


knitr::kable(tab_gender_display, caption = tr("table.respondents_by_gender"))

cat("### ", tr("table.age_group"), "\n\n")
knitr::kable(
  tab_age_group_display,
  caption = tr("table.respondents_by_age_group")
)

cat("### ", tr("table.geo"), "\n\n")

knitr::kable(
  tab_geo_display,
  caption = paste(
  tr_variable(adm1, "short"),
  tr("common.and"),
  tr_variable(adm2, "short")
)
)

```

## `r tr("report.weighting")`

```{r weighting, echo=FALSE, results='asis'}

include_text("weighting")

weighting_plot

```

## `r tr("report.correlation")`

```{r correlation, results='asis',fig.width=corr_plot_height}

if (module == "inst") {
corrplot(cor_matrix_w, p.mat = testRes$p, sig.level = 0.10, method = 'circle', type = 'lower', order = 'hclust', addrect = 2 , col = COL2('PuOr', 10),insig='blank',
         tl.col = color_label_grey, tl.cex = 0.9,
         addCoef.col = "white", number.cex = 0.7, diag = FALSE, addgrid.col = 'white')


include_text("correlation")

} else if (is.function(correlation_ews_plot)) {
  correlation_ews_plot()
  
  include_text("correlation")
  
} else {}
```

## `r tr("report.significance_testing")`

```{r significance_testing, echo=FALSE,results='asis', fig.width=7, fig.height=3}


include_text("significance")


## Significant testing table

tab_sig_reshaped<-formattable(sig_reshaped, align =c("c","c", "c", "c", "c"))
tab_sig_reshaped
```

# `r tr("report.download")`

```{r download_button, echo=FALSE, results='asis'}

# ============================================================
# Export data
# ============================================================


source(paste0(repo,"/R/base/export.R")) # export to excel and supabase


cat(
  paste0(
    '::: div-button\n',
    '[',tr("report.download_button"),'](', export_file, '){.button}\n',
    ':::\n'
  )
)
```
